### 2016-09-16 image_setup.tex
changed fdisks
Corrections

The latest Raspbian Lite bundle can be downloaded from the Raspbian
official webpage~\cite{raspbian}. At the time of this writing, the latest
available bundle was \texttt{2016-05-27-raspbian-jessie-lite.zip}.
To ensure that the content of the bundle does not change, this procedure
is based on that particular version of Raspbian Lite which we have
made available at~\cite{tunescode_webpage} acting as our \ac{HTTP} server.
All other files used in this paper are available there unless specified
otherwise. To get started, the testbed administrator must open a Linux
shell (terminal) in its \ac{PC} and declare the environment variables shown
in the command block below. We show the whole procedure by performing the role
of the testbed administrator.
%To copy the commands from this procedure to the
%testbed administrator terminal, we recommend to read this document
%with Adobe Acrobat Reader in Linux. Otherwise, copied characters might not
%render properly when pasted.

@@

Now, in this command block output, we observe that the change has taken
effect by noticing the total available image size is 2.3 GiB. To expand the
root filesystem, it is required to first remove the Linux partition
and then add it again with one GB more. To do this, we make use
of \texttt{fdisk} again in command mode to alter the partition table.
We pass the arguments of the command mode of \texttt{fdisk}
through the \texttt{echo} command in Linux and the \texttt{|} operator
(pipe) as follows:

\begin{lstlisting}[]
$ (echo d; echo 2; echo n; echo ; echo ; echo 137216; echo ; echo w) | fdisk ${IMAGE}.img
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

The \texttt{echo} commands within the parenthesis are interpreted as
key-presses in the \texttt{fdisk} commmand mode. They (i) delete partition
number 2, (ii) create a new primary type partition by default, (iii) set the
new partition starting point (137216) and (iv) write new partition table to
the image file. The previously shown command options of \texttt{fdisk}
should work for both old and new versions of the commands. However, if this
is not the case, the testbed administrator can do these commands manually.
In this case, it is only required to delete the old root
filesystem partition and create the new one starting from the same sector
number. In our case, if the partitions commands were correct, the output
should be following:

\begin{lstlisting}[]
Command (m for help): Partition number (1-4):
Command (m for help): Partition type:
   p   primary (1 primary, 0 extended, 3 free)
   e   extended
Select (default p): Using default response p
Partition number (1-4, default 2): Using default value 2
First sector (2048-4806655, default 2048): Last sector, +sectors or
+size{K,M,G} (137216-4806655, default 4806655): Using default value 4806655

Command (m for help): The partition table has been altered!

Syncing disks.

\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

\subsection{Image OS Files and Configuration Scripts Setup}
In general, the \ac{Raspi}s should be setup as similar as possible. However,
some particularities exist to differentiate the devices in principle. Also,
scripts containing further configurations for the \ac{Raspi}s are desirable
to be distributed as part of the common image. Therefore, here we present
the steps to setup basic properties of the \ac{Raspi}s and distributing
configuration scripts to each of them through the image. For this, we first
indicate how to obtain and put our configuration scripts in the image. Later,
we describe the tasks performed by these configuration scripts. Finally, we
indicate how and in which order the scripts are executed to configure all the
devices. Any testbed administrator might modify or include other
tasks besides the presented in this procedure according to his / her needs
as we will show.

@@

The unzippped files are one configuration file and three configuration
scripts put in the newly created \texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/}
folder in the image. In what follows, we describe which features do we
require all the \ac{Raspi}s to have and how are they achieved with
these configuration scripts.

\subsubsection{Device Hostnames}
Among the settings that we want to be different in the devices is
their hostname. The hostname helps the user to physically distinguish the
devices from each other. Thus, in our case we require the devices in our
testbed to have different hostnames. We define the hostnames based
on the \ac{MAC} addresses of the \ac{Raspi}s wired Ethernet interface.

@@

The testbed administrator has to insert the \ac{MAC} addresses and hostnames
of his / her \ac{Raspi}s obtained previously in the format shown in the
configuration file. In that file format, on each line there is a \ac{MAC}
address and the corresponding hostname for the given \ac{Raspi}. For
simplicity in the example above, we just include four \ac{Raspi}s.
This file will be employed by the
\texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/set\_hostname}
\ac{Bash} script to assign the hostname of each \ac{Raspi}.
The script content is the following:

>>>>

The latest Raspbian Lite bundle can be downloaded from the Raspbian
official webpage~\cite{raspbian}. At the time of this writing, the latest
available bundle was \texttt{2016-05-27-raspbian-jessie-lite.zip}.
To ensure that the content of the bundle does not change, this procedure
is based on that particular version of Raspbian Lite which we have
made available at~\cite{soerensen_chres_wiant_2016_154143}. All other files used in this paper are available there. The testbed
administrator has to move these files to his / her \ac{HTTP} server.
To get started, the testbed administrator must open a Linux
shell (terminal) in its \ac{PC} and declare the environment variables shown
in the command block below. We show the whole procedure by performing the role
of the testbed administrator.
%To copy the commands from this procedure to the
%testbed administrator terminal, we recommend to read this document
%with Adobe Acrobat Reader in Linux. Otherwise, copied characters might not
%render properly when pasted.

@@

Now, in the above command block output, we observe that the change has taken
effect by noticing the total available image size is 2.3 GiB. To expand the
root filesystem, we replace the Linux partition with a new partition one GB
larger. The starting point of this new partition should be the same as the old
one. We make use of \texttt{fdisk} to alter the partition table
in the commands below. They (i) delete partition number 2, (ii) create a new
primary partition and (iii) set the new partition starting point. The new
partition starting point value is 137216 in our case. Finally, we (iv) write
the new partition table to the image file. This is made as follows:

%\begin{lstlisting}[]
%$ (echo d; echo 2; echo n; echo ; echo ; echo 137216; echo ; echo w) | fdisk ${IMAGE}.img
%\end{lstlisting}
\begin{lstlisting}[]
$ fdisk ${IMAGE}.img << EOF
d
2
n
p
2
137216

w
EOF
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

If the partitions commands were correct, the partition table should now
look like the following: \todot{PRINT PARTITION TABLE}

% \begin{lstlisting}[]
% Command (m for help): Partition number (1-4):
% Command (m for help): Partition type:
%    p   primary (1 primary, 0 extended, 3 free)
%    e   extended
% Select (default p): Using default response p
% Partition number (1-4, default 2): Using default value 2
% First sector (2048-4806655, default 2048): Last sector, +sectors or
% +size{K,M,G} (137216-4806655, default 4806655): Using default value 4806655

% Command (m for help): The partition table has been altered!

% Syncing disks.

% \end{lstlisting}
% \FloatBarrier
% \vspace{-5mm}

@@

\subsection{Image OS Files and Configuration Scripts Setup}
In general, the \ac{Raspi}s should be setup as similar as possible. However,
some particularities exist to differentiate the devices in principle. Also,
scripts containing further configurations for the \ac{Raspi}s are desirable
to be distributed as part of the common image. Therefore, here we present
the steps to setup basic properties of the \ac{Raspi}s and distributing
configuration scripts to each of them through the image. For this, we first
indicate how to obtain and put our configuration scripts in the image. Later,
we describe the tasks performed by these configuration scripts. Finally, we
indicate how and in which order the scripts are executed to configure all the
devices. Any testbed administrator might modify or include other
tasks according to his / her needs as we will show.

@@

The unzippped files are one configuration file and three configuration
scripts in the newly created \texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/}
folder in the image. In what follows, we describe which features do we
require all the \ac{Raspi}s to have and how are they achieved with
these configuration scripts.

\subsubsection{Device Hostnames}
The hostname helps the user to physically distinguish the
devices from each other. In our case we require the devices in our
testbed to have different hostnames. We define the hostnames based
on the \ac{MAC} addresses of the \ac{Raspi}s wired Ethernet interface.

@@

The testbed administrator has to insert the \ac{MAC} addresses and hostnames
of his / her \ac{Raspi}s obtained previously in the format shown in the
configuration file. For each given \ac{Raspi}, there is a \ac{MAC} address and
the corresponding hostname. This file will be employed by the
\texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/set\_hostname} \ac{Bash} script to
assign the hostname of each \ac{Raspi}. The script content is the following:

### 2016-09-16 tools.tex
Corrections

%\subsection{Kodo cross-compilation: From your PC to the Raspberry Pi}
\label{sec:tools}
Within the testbed everyday use, there exists frequent tasks that require
a set of various commands in a given \ac{Raspi}. This could be tedious, prone
to errors and time consuming to realize everytime the task is required to be
made. Therefore, in this section we introduce a set of tools that help to
automate and monitor routinary tasks execution in the \ac{Raspi}s and show
relevant example commands with them. To be able to run all the following
commands, it is required to have \ac{SSH} connectivity with the \ac{Raspi}s,
otherwise the commands need to be run locally in a \ac{Raspi} making
necessary to use a keyboard and a monitor. The testbed
administrator needs to put the memory cards in the \ac{Raspi}s and turn them
on, for them to be able to boot. The devices should boot properly and have
connectivity if the image and \ac{DHCP} server were properly set previously.

\subsection{Fabric}
Controlling multiple devices using \ac{SSH} from a single \ac{PC}
often leads to many repetitive tasks. Among these, we can mention:
(i) rebooting a set of devices, (ii) installing applications in multiple
devices and (iii) copying files to and / or from multiple devices and so on. To
simplify the management and avoid execution issues, Fabric~\cite{python_fabric} provides
a Python library to ease the major task of working with multiple devices from
a single \ac{PC}. First, the testbed administrator creates a directory to
hold the \texttt{Fabric} source code:

>>>>

%\subsection{Kodo cross-compilation: From your PC to the Raspberry Pi}
\label{sec:tools}
Within the testbed everyday use, there exists frequent tasks that require
a set of various commands in a given \ac{Raspi}. This could be tedious, prone
to errors and time consuming to realize everytime the task is required to be
made. Therefore, in this section we introduce a set of tools that help to
automate and monitor routinary tasks execution in the \ac{Raspi}s and show
relevant example commands with them. To be able to run all the following
commands, it is required to have \ac{SSH} connectivity with the \ac{Raspi}s,
otherwise the commands need to be run locally on a \ac{Raspi} making
necessary to use a keyboard and a monitor. The testbed
administrator needs to put the memory cards in the \ac{Raspi}s and turn them
on, for them to be able to boot. The devices should now be bootable.

\subsection{Fabric}
Controlling multiple devices using \ac{SSH} from a single \ac{PC}
often leads to many repetitive tasks. Among these, we can mention:
(i) rebooting a set of devices, (ii) installing applications in multiple
devices and (iii) copying files to / from multiple devices.
Fabric~\cite{python_fabric} provides a Python library that simplifies the
management of working with many devices from a single \ac{PC}. First, the
testbed administrator creates a directory to hold the \texttt{Fabric} source
code:
